Group assignment

### 2022-11-05 Draft.Rmd
Update Draft.Rmd

---
title: "DRAFT"
author: "Alima Dzhanybaeva"
date: "2022-10-31"
output:
  pdf_document:
         latex_engine: xelatex
  html_document: default
---
```{r include=FALSE}
library(tidyverse)
library(modelsummary)
library(dplyr)
library(knitr)
library(htmlTable)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
## 1.1 Overview of the original data.
### The original hotels-europe data includes information on price and features of hotels in 46 European cities and for 2017-2018. The data was downloaded from https://osf.io/yzntm/download.

@@

### • The folowing manipulations were carried out on the original data to change the existing variables and to add new ones:
* The type of the variables *center1distance* and *guestreviewsrating* were changed from 'string' to 'numeric'. The new variable names are *distance* and *actualrating * respectively.
```{r include=FALSE}
# guestreviewrating separation
df2 <- separate(df2, guestreviewsrating, ' /', into =
                 c('actualrating', 'rate'))
df2 <- select(df2, -rate)
df2$actualrating <- as.numeric(df2$actualrating)

@@

df2$country <- as.factor(df2$addresscountryname)
df2 <- select(df2, -addresscountryname)
```
* The original *accommodationtype* variable was tranformed into new *acctype_f* factor variable.
```{r}
df2 <- separate(df2, accommodationtype, '@', into = 
                 c('word', 'acctype'))
df2 <- select(df2, -word)
df2 <- mutate(df2, acctype_f = factor(acctype))
```
* *trueprice* variable was generated by dividing the original *price* variable by *price_night*
```{r}
df2 <- separate(df2, price_night, ' ', into =
                     c('pr','word', 'nights', 'night'))
df2 <- select(df2, -pr)
df2 <- select(df2, -night)
df2 <- select(df2, -word)
df2$nights <- as.numeric(df2$nights)
df2$trueprice <- df2$price/df2$nights
```
* *season* variable was created based on the *month* variable
```{r}
df2 <- df2 %>%
  mutate(season = case_when(month == 12 | month == 1 |month == 2 ~ 'winter',
                            month == 3 | month == 4 |month == 5 ~ 'spring',
                            month == 6 | month == 7 |month == 8 ~ 'summer',
                            month == 9 | month == 10 |month == 11 ~ 'autumn'))
df2$season <- as.factor(df2$season)
```
* With the help of the existing *starrating* variable new *class* variable was generated
```{r}
df2 <- df2 %>%
  mutate(class = case_when(starrating == 1.0 | starrating == 1.5 ~ 'tourist',
                           starrating == 2.0 | starrating == 2.5 ~ 'standard',
                           starrating == 3.0 | starrating == 3.5 ~ 'comfort',
                           starrating == 4.0 | starrating == 4.5 ~ 'first class',
                           starrating == 5.0 ~ 'luxury'))
df2$class <- as.factor(df2$class)
```

@@

## 1.3 Findings. 
### For this project, we wanted to find the best deal among different types of accommodation in Central European countries. 
### Consequently, we formulated the following question: 
* *How does the average price for accommodation change for different values of the following variables: year, country, distance, season, class?*

### • How the price for accommodation changed for 7 Central European countries in 2018 in comparison with 2017.
```{r}
df3 <- aggregate(trueprice ~ country + year, 
                 data=df2, 
                 function(x) { 
                   c(mean_price = mean(x)) 
                 })
ggplot(data = df3, aes(x=year, y=trueprice, fill=country)) + 
  geom_bar(stat='identity') + facet_wrap(~country) +
  geom_text(aes(label = round(trueprice)), vjust = -0.2) +
  scale_x_discrete() + ylim(0, 195) + 
  labs(x='Year(2017,2018)', y='Price for one night (in EUR)', fill='Country')
```

### As we can see from the graph, the price increased for 6 countires, Slovakia being the only case in which the price dropped.
### The biggest jump in price was in Italy. In this country, the price in 2018 increased by 24 euros compared to 2017.

### • Mean prices for countries

@@

### • Mean price/ distance

@@

### • Mean prices for diff seasons
```{r echo=FALSE}
datasummary(trueprice*season ~ Mean, data=df2) %>%
   kableExtra::kable_styling(latex_options = "hold_position")
```

### • Plot for classes
```{r echo=FALSE}
datasummary(trueprice*class ~ Mean, data = df2) %>%
   kableExtra::kable_styling(latex_options = "hold_position")
```
```{r echo=FALSE, message=FALSE}
ggplot(data = df2, aes(x=reorder(class, +trueprice), y=trueprice, fill = class)) + 
  geom_bar(stat = 'summary') + labs(x='Class', y='Price for 1 night (in EUR)', fill = 'Class')
```

### Weekend
```{r}
ggplot(data=df2, aes(x=weekend, y=trueprice, group=weekend))+
  geom_violin(trim=FALSE, fill='#A4A4A4', color="darkred")+
  geom_boxplot(width=0.1) + theme_minimal() + 
  labs(x="Weekend", y="Price for one night (in EUR)")
```

>>>>

---
title: "Group assignment"
author: "Alima Dzhanybaeva, Jana Hochel, Tasnim Murad"
date: "2022-05-11"
output:
  pdf_document:
         latex_engine: xelatex
  html_document: default
---
```{r include=FALSE}
library(tidyverse)
library(modelsummary)
library(dplyr)
library(knitr)
library(htmlTable)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
## 1.1 Overview of the original data.
### The original hotels-europe data includes information on price and features of hotels in 46 European cities and for 2017-2018. The data was downloaded from https://osf.io/yzntm/download.

@@

### • The folowing manipulations were carried out on the original data to change the existing variables and to add new ones:
* The type of the variables *center1distance* and *guestreviewsrating* were changed from 'string' to 'numeric', and from 'string' to 'factor' for *addresscountryname*. The new variable names are *distance*, *actualrating * and *country* respectively.
```{r include=FALSE}
# guestreviewrating separation
df2 <- separate(df2, guestreviewsrating, ' /', into =
                 c('actualrating', 'rate'))
df2 <- select(df2, -rate)
df2$actualrating <- as.numeric(df2$actualrating)

@@

df2$country <- as.factor(df2$addresscountryname)
df2 <- select(df2, -addresscountryname)
```
* The original *accommodationtype* variable was transformed into new *acctype_f* factor variable.
```{r}
df2 <- separate(df2, accommodationtype, '@', into = 
                 c('word', 'acctype'))
df2 <- select(df2, -word)
df2 <- mutate(df2, acctype_f = factor(acctype))
```
* *trueprice* variable was generated by dividing the original *price* variable by *price_night*
```{r}
df2 <- separate(df2, price_night, ' ', into =
                     c('pr','word', 'nights', 'night'))
df2 <- select(df2, -pr)
df2 <- select(df2, -night)
df2 <- select(df2, -word)
df2$nights <- as.numeric(df2$nights)
df2$trueprice <- df2$price/df2$nights
```
* *season* variable was created based on the *month* variable
```{r}
df2 <- df2 %>%
  mutate(season = case_when(month == 12 | month == 1 |month == 2 ~ 'winter',
                            month == 3 | month == 4 |month == 5 ~ 'spring',
                            month == 6 | month == 7 |month == 8 ~ 'summer',
                            month == 9 | month == 10 |month == 11 ~ 'autumn'))
df2$season <- as.factor(df2$season)
```
* With the help of the existing *starrating* variable new *class* variable was generated
```{r}
df2 <- df2 %>%
  mutate(class = case_when(starrating == 1.0 | starrating == 1.5 ~ 'tourist',
                           starrating == 2.0 | starrating == 2.5 ~ 'standard',
                           starrating == 3.0 | starrating == 3.5 ~ 'comfort',
                           starrating == 4.0 | starrating == 4.5 ~ 'first class',
                           starrating == 5.0 ~ 'luxury'))
df2$class <- as.factor(df2$class)
```

@@

## 1.3 Analysis 
### For this project, we wanted to find the best deal among different types of accommodation in Central European countries. 
### Consequently, we formulated the following question: 
* *How does the average price for accommodation change for different values of the following variables: year, country, distance, season, class, weekend?*

## *Findings*

### • As we can see from the graph below, the price increased for 6 countires, Slovakia being the only case in which the price dropped.
### The biggest jump in price was in Italy. In this country, the price in 2018 increased by 24 euros compared to 2017.
```{r}
df3 <- aggregate(trueprice ~ country + year, 
                 data=df2, 
                 function(x) { 
                   c(mean_price = mean(x)) 
                 })
ggplot(data = df3, aes(x=year, y=trueprice, fill=country)) + 
  geom_bar(stat='identity') + facet_wrap(~country) +
  geom_text(aes(label = round(trueprice)), vjust = -0.2) +
  scale_x_discrete() + ylim(0, 195) + 
  labs(x='Year(2017,2018)', y='Price for one night (in EUR)', fill='Country')
```

### • The lowest price is 12 EUR per night (Czech Republic),  whereas the lowest average cost is 86 EUR (Poland). On average, Austrian accommodation is the most expensive with a mean price of 158 EUR.

@@

### • The eminent negative relationship between price and distance from the city center can be tracked only until 7.5 miles. After that point, the dependence of the accommodation cost on the *distance* variable becomes vaguer, as the points are much more dispersed and do not follow a specific pattern. Therefore, we can conclude that on average, the cheapest accommodation is 7.5 miles far from the city center.

@@

### • The prices for accommodation in the chosen 7 Central European countries are on average the lowest during autumn and are highest during spring. The difference between the lowest and the highest value is 35 euros.

```{r echo=FALSE}
datasummary(trueprice*season ~ Mean, data=df2) %>%
   kableExtra::kable_styling(latex_options = "hold_position")
```

### • As expected, the cheapest class is "tourist" (1 and 1.5 stars) with a mean price of 88 euros. However, the average cost for "standard" (2 and 2.5 stars) falls not that far apart from the best deal with the difference of only 5 euros. Conversely, the "luxury"  (5 stars) class is on average 140 euros more expensive than  "first class" (4 and 4.5 stars). 
### Consequently, the difference between mean prices becomes bigger with the number of stars.

```{r echo=FALSE}
datasummary(trueprice*class ~ Mean, data = df2) %>%
   kableExtra::kable_styling(latex_options = "hold_position")
```
```{r echo=FALSE, message=FALSE}
ggplot(data = df2, aes(x=reorder(class, +trueprice), y=trueprice, fill = class)) + 
  geom_bar(stat = 'summary') + labs(x='Class', y='Price for 1 night (in EUR)', fill = 'Class')
```

### • The average price for accommodation is 16 euros higher during weekends. Moreover, as we can see from the graph below: (1) the boxplot for weekends is comparatively taller than the one for weekdays, indicating that prices are more dispersed during the end of the week; (2) the upper quartile for weekends is farther from the median than for weekdays, meaning that the 75% percentile is higher.

```{r echo=FALSE}
df2 <- df2 %>%
  mutate(weekend = case_when(weekend == 1 ~'weekend',
                            weekend == 0 ~ 'weekday'))
datasummary(trueprice*weekend ~ Mean, data = df2) %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```
```{r echo=FALSE}
ggplot(data=df2, aes(x=weekend, y=trueprice, group=weekend))+
  geom_violin(trim=FALSE, fill='#A4A4A4', color="darkred")+
  geom_boxplot(width=0.1) + theme_minimal() + 
  labs(x="Weekend", y="Price for one night (in EUR)")
```

## *Summary*
### For the chosen 7 Central European countries (with the maximum value for the *distance* variable set to 35 miles and to 600 euros for *trueprice* variable)  following conclusions on the best deal accross different accommodation types were made:
* Compared to other countries, the average prices in Slovakia do not tend to drastically jump.
* On average, the cheapest deals can be found in Poland.
* The optimal distance from the city center is 7,5 miles.
* Autumn is the best season for finding advantageous offers.
* "Tourist" is the class with the lowest prices, however, "standard" is not much more expensive.
* Accommodation on weekdays is not only cheaper, but the prices are also not as dispersed as on weekends.
